---
title: "The Association Between Emotion Regulation Strategies and Depressive Symptoms among Adolescent Girls"
authors: "Xinyu (Lily) Wang, Sophie Hartford, Everett Mahaffy"
format:
  html:
    toc: true
    embed-resources: true
    highlight-style: tango
execute:
  eval: true
  echo: false
warning: false
message: false
bibliography: references.bib
csl: apa.csl
---

# Abstract

Adolescence is a developmental period marked by heightened emotional reactivity and increasing vulnerability to depression, particularly among girls. Emotion regulation (ER) strategies such as cognitive reappraisal and expressive suppression are known correlates of mental health in older adolescents and adults. Yet, less is known about how these strategies relate to depressive symptoms during early adolescence. The present study examined associations between ER strategies and depressive symptoms in a sample of 174 adolescent girls aged 10-13 from the Transitions in Adolescent Girls (TAG) study. Linear regression models showed that greater use of cognitive reappraisal was associated with lower depressive symptoms, whereas greater use of expressive suppression was associated with higher depressive symptoms. These effects remained significant in a multiple regression model including both ER strategies. Exploratory moderation analyses revealed that age did not moderate the association between reappraisal and depressive symptoms or suppression and depression. Findings highlight early adolescence as a critical window for promoting adaptive ER skills and reducing depression risk, underscoring the importance of early prevention and intervention efforts.

*Keywords: Adolescent depression, emotion regulation, developmental timing*

```{r}

# Load packages

library(tidyverse)
library(flextable)
library(here)
library(readxl)
library(modelsummary) # for tables
library(janitor) # for cleaning names
library(performance) # for assessing regression assumptions
```

```{r}

# Load data

# Demographic
age <- read_csv(here("data", "TAG_age_session_dates_4waves.csv"))
age <- clean_names(age)

race <- read_excel(here("data", "TAG_W1_Race_Ethnicity.xlsx"))
race <- clean_names(race)

# Emotion Regulation(ERQ)
erq <- read_csv(here("data", "ERQ_Wave1.csv"))
erq <- erq %>% 
       clean_names() %>% 
       distinct(tagid, .keep_all = TRUE)

# Adolescent Depression(CESDC)
cesdc <- read_csv(here("data", "CESDC_Wave1.csv"))
cesdc <- clean_names(cesdc)
```

```{r}

# Participant demographics
## Race/ethnicity category
### Clean the data frame: identify subjects reported multiple racial categories and rename them to "Multi-racial"

race_clean <- race %>%
  mutate(
    w1_ethnicity = as.character(w1_ethnicity),
    w1_ethnicity = if_else(
      str_detect(w1_ethnicity, ","),
      "g. Multi-racial",
      w1_ethnicity
    )
  )
```

```{r}

# Age
## We have subject age data at both sessions and we are interested in calculating the mean age of the two sessions of wave1. 

age <- age %>% 
  mutate(W1_mean_age = (w1s1_age + w1s2_age)/2) 

# Check mean age for sample
age_summary <- age %>%
  summarize(
    mean_W1_age = mean(W1_mean_age, na.rm = TRUE),
    sd_W1_age   = sd(W1_mean_age, na.rm = TRUE),
  )
#print(age_summary)
```

```{r}
# Join all data frames

df <- erq %>% 
  left_join(cesdc, by= 'tagid')

df <- df %>% 
     left_join(age, by = 'tagid')

df <- df %>% 
     left_join(race_clean, by = 'tagid') %>% 
     select(tagid, erq_reappraisal_total, erq_suppression_total, ces_dc_total_75perc, W1_mean_age, w1_ethnicity, 
            w1s1_date, w1s1_age, w1s2_date, w1s2_age, w2s1_date, w2s1_age, w2s2_date, w2s2_age, 
            w3s1_date, w3s1_age, w3s2_date, w3s2_age, w4s1_date, w4s1_age, w4s2_date, w4s2_age
            )

df <- df %>%
  rename(
    wave1_session1_date = w1s1_date,
    wave1_session1_age  = w1s1_age,
    wave1_session2_date = w1s2_date,
    wave1_session2_age  = w1s2_age,

    wave2_session1_date = w2s1_date,
    wave2_session1_age  = w2s1_age,
    wave2_session2_date = w2s2_date,
    wave2_session2_age  = w2s2_age,

    wave3_session1_date = w3s1_date,
    wave3_session1_age  = w3s1_age,
    wave3_session2_date = w3s2_date,
    wave3_session2_age  = w3s2_age,

    wave4_session1_date = w4s1_date,
    wave4_session1_age  = w4s1_age,
    wave4_session2_date = w4s2_date,
    wave4_session2_age  = w4s2_age
  )
```

```{r}
# Tidy
df2 <- df %>%
  mutate(across(starts_with("wave"), as.character))

df_long <- df2 %>%
  pivot_longer(
    cols = starts_with("wave"),
    names_to = c("wave", "session", "info"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  mutate(
    wave    = parse_number(wave),
    session = parse_number(session)
  )

df_tidy <- df_long %>%
  pivot_wider(
    names_from = info,
    values_from = value
  ) %>% 
  mutate(age = as.numeric(age))


# Filter
df_final <- df_tidy %>% 
  filter(wave == 1, session == 1)

```

# Introduction

## Depression in Adolescent Girls

Major depressive disorder (MDD) and depressive symptoms are among the most significant mental health concerns affecting adolescents [@benton2021]. Measured in Years Lived with Disability (YLD), depression is one of the leading causes of disability worldwide, contributing significantly to the global burden of disease [@reddy2010]. @IHME2019 reported that depression had become the second leading cause of disability, with 605.67 YLDs per 100,000 in 2019.

The issue has become critical among adolescents. According to the @NIMH_MajorDepression, the prevalence of MDD among adolescents reached 5 million, accounting for 20.1% of the U.S. population aged 12-17. The prevalence was higher among adolescent females (29.2%) compared to adolescent males (11.5%). MDD is significantly associated with a range of developmental outcomes, such as educational attainment [@breslau2008; @breslau2011; @lee2009], early childbearing [@marcus2009], and occupational functioning and age at marriage [@fakhari2020], implying that MDD is associated with adverse health outcomes and long-term impairment, particularly when it begins in adolescence. Given its high prevalence and far-reaching consequences, it is critical to identify the mechanisms, as well as the risk and resilience factors, that contribute to the onset and progression of depressive symptoms in adolescence.

## Emotion Regulation Strategies

Although adolescents are, on average, more often happy than unhappy [@larson2002; @mclaughlin2015], several critical emotional changes occur during adolescence [@coe-odess2019]. Adolescents show greater variability in daily affect [@larson2002; @maciejewski2017], higher levels of negative affect, and lower levels of positive affect compared to younger children [@abitante2022; @griffith2021]. These changes in affect likely place increased demands on adolescents’ ability to manage such emotions, underscoring the importance of effective ER skills in adolescence. These shifts appear to occur between early and middle adolescence, highlighting a potential sensitive period for intervention.

ER strategies differ markedly in their effectiveness. Strategies such as reappraisal, acceptance, and distraction reliably reduce negative affect and enhance positive affect [@boemo2022; @wante2018], whereas rumination, expressive suppression, and worry are typically linked to increases in negative affect [@boemo2022]. Among these, cognitive reappraisal and expressive suppression are two of the most widely studied strategies, and both have been strongly implicated in mental health outcomes. Cognitive reappraisal involves reframing an emotional stimulus and redirecting attention to the emotional scenario to change its emotional impact [@gross2015; @gross2003]. While expressive suppression involves the inhibition of overt expression of emotion [@gross2015; @gross2003].

## Emotion Regulation and Mental Health

ER has been identified as a transdiagnostic factor in both internalizing and externalizing psychopathology during adolescence [@aldao2016; @özlemschäfer2017; @joormann2016]. Cognitive emotion regulation strategies (e.g., cognitive reappraisal) appear to be particularly salient predictors of later mental health in adolescence [@tebrinke2021; @zagaria2023]. Thus, ER has been proposed as a key early identification factor for detecting mental health concerns [@murray2024]. Difficulties with ER, like emotion maladaptation, are strong predictors and early signs of depression, especially in youth [@paulus2021]. However, the majority of work investigating the relation between ER and depression has focused on adult or late-adolescent populations. Given that early adolescence has been proposed as a particularly pivotal period [@silvers2022; @uhlhaas2023], the present study sought to examine these associations in early adolescence, focusing on girls given their elevated risk of developing depression.

## Aims and Hypotheses

We aimed to examine the association between emotion regulation strategies (reappraisal and suppression) and depressive symptoms among young adolescent girls, and we predicted 1). higher reappraisal usage will be associated with lower depressive symptoms, and 2). higher suppression usage will be associated with higher depressive symptoms. Additionally, exploratory analyses examined whether each ER strategy remained associated with depressive symptoms when adjusting for the other strategy, and whether age moderated these associations.

# Method

## Participants

The present study uses baseline data from the Transitions in Adolescent Girls (TAG) study, a longitudinal project examining the developmental trajectories of pubertal, neurological, and social processes as mechanisms underlying the emergence of mental illness in female adolescents [@barendse2019]. A total of 174 adolescent girls were recruited from the local community in Lane County, Oregon, USA, primarily through letters sent by schools to all families with children registered as female students. The inclusion criteria required participants to be 10-13 years old at enrollment, fluent in English, and have normal or corrected-to-normal vision. Exclusion criteria included current use of psychotropic medication, MRI contraindications (e.g., claustrophobia, presence of ferromagnetic material in the body), suspected or confirmed pregnancy, and diagnoses of a developmental disability, psychotic disorder, or behavioral disorder (including autism). Parents or legal guardians provided informed consent, and adolescents provided assent before participation.

### Figure 1. Plot of TAG Participants

```{r}

ggplot(df_tidy, aes(x = age, y = tagid)) +
  geom_line(aes(group = tagid), size = .5, alpha = .4) +
  geom_point(aes(color = as.factor(wave)), size = 2) +
  labs(
    x = "Age",
    y = "TAG Participant",
    title = "Age by Wave",
    color = "Wave"
  ) +
  theme_classic() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
 scale_x_continuous(breaks = seq(10, 20, by = 0.5))

```

## Measures

### Adolescent Depressive Symptoms

Depressive symptoms were assessed using the Center for Epidemiological Studies Depression Scale for Children (CES-DC), a 20-item self-report instrument designed to measure the severity of depressive symptoms over the past week. Each item is rated on a 4-point Likert scale, ranging from 0 (“Not at all”) to 4 (“A lot”), with higher total scores indicating greater levels of depressive symptomatology [@weissman1980] . A score of 15 (out of 60) is considered indicative of clinically significant depressive symptoms.

### Emotion Regulation Strategies

Emotion regulation strategies were measured using the Emotion Regulation Questionnaire (ERQ) [@gross2003]. The ERQ contains 10 items that assess individuals’ use of emotion regulation, with six items assessing cognitive reappraisal and four assessing expressive suppression. Each item is rated on a 7-point Likert scale, ranging from 1 (“strongly disagree”) to 7 (“strongly agree”). The present analyses treat each scale (reappraisal and suppression) separately.

## Analytic Plan

All analyses were conducted in *R* [@R]. Data preparation utilized the following packages: *tidyverse* [@R-tidyverse], *here* [@R-here], *readxl* [@R-readxl], *janitor* [@R-janitor], *modelsummary* [@R-modelsummary], *performance* [@R-performance], and *flextable* [@R-flextable]. For primary research questions 1 and 2, we estimated linear regression models to quantify the relation between the use of each ER strategy and depressive symptoms. For exploratory research question 1, we estimated a linear multiple regression model, which included both reappraisal and suppression as predictors. For exploratory research questions 2 and 3, we included an interaction term to explore adolescent age as a potential moderator. All *p*-values were evaluated at *α* = .05.

# Results

## Descriptive Statistics

Sample descriptive statistics are reported in Table 1. Participants were young adolescent girls with a mean age of `r round(mean(df_final$W1_mean_age, na.rm = TRUE), 1)` years (SD = `r round(sd(df_final$W1_mean_age, na.rm = TRUE), 1)`, range = `r round(min(df_final$W1_mean_age, na.rm = TRUE), 1)`–`r round(max(df_final$W1_mean_age, na.rm = TRUE), 1)`). On average, adolescents reported moderate use of emotion regulation strategies. Mean reappraisal scores were `r round(mean(df_final$erq_reappraisal_total, na.rm = TRUE), 1)` (SD = `r round(sd(df_final$erq_reappraisal_total, na.rm = TRUE), 1)`, range = `r round(min(df_final$erq_reappraisal_total, na.rm = TRUE), 1)`–`r round(max(df_final$erq_reappraisal_total, na.rm = TRUE), 1)`), and mean suppression scores were `r round(mean(df_final$erq_suppression_total, na.rm = TRUE), 1)` (SD = `r round(sd(df_final$erq_suppression_total, na.rm = TRUE), 1)`, range = `r round(min(df_final$erq_suppression_total, na.rm = TRUE), 1)`–`r round(max(df_final$erq_suppression_total, na.rm = TRUE), 1)`). Depressive symptoms were generally low to moderate, with a mean score of `r round(mean(df_final$ces_dc_total_75perc, na.rm = TRUE), 1)` (SD = `r round(sd(df_final$ces_dc_total_75perc, na.rm = TRUE), 1)`, range = `r round(min(df_final$ces_dc_total_75perc, na.rm = TRUE), 1)`–`r round(max(df_final$ces_dc_total_75perc, na.rm = TRUE), 1)`).

The racial/ethnic composition of the sample was predominantly White/Caucasian (`r round(mean(df_final$w1_ethnicity == "d. White / Caucasian", na.rm = TRUE) * 100, 1)`%), followed by Multiracial (`r round(mean(df_final$w1_ethnicity == "g. Multi-racial", na.rm = TRUE) * 100, 1)`%) and Hispanic/Latino/Chicano (`r round(mean(df_final$w1_ethnicity == "b. Hispanic/ Latino/ Chicano", na.rm = TRUE) * 100, 1)`%). Smaller proportions of participants identified as Asian (`r round(mean(df_final$w1_ethnicity == "e. Asian", na.rm = TRUE) * 100, 1)`%), Black/African American (`r round(mean(df_final$w1_ethnicity == "a. Black/ African American", na.rm = TRUE) * 100, 1)`%), Native American/Alaska Native (`r round(mean(df_final$w1_ethnicity == "c. Native American or Native Alaskan", na.rm = TRUE) * 100, 1)`%), or Other (`r round(mean(df_final$w1_ethnicity == "h. Other", na.rm = TRUE) * 100, 1)`%). A small percentage declined to respond (`r round(mean(df_final$w1_ethnicity == "j. Decline to respond", na.rm = TRUE) * 100, 1)`%) or had missing ethnicity data (`r round(mean(is.na(df_final$w1_ethnicity)) * 100, 1)`%).

### Table 1. Sample Summary Statistics

```{r}

# Sample Summary Statistics

df_names <- df_tidy %>% 
    select(-wave, -session, -age) %>% 
    rename(
    Reappraisal       = erq_reappraisal_total,
    Suppression       = erq_suppression_total,
    Depression        = ces_dc_total_75perc,
    Age               = W1_mean_age,
    Ethnicity         = w1_ethnicity
    )
datasummary_skim(df_names)
```

```{r}

# RQ 1

## Is higher emotion reappraisal associated with lower depression symptoms in adolescent girls?

rq1<-lm(ces_dc_total_75perc~erq_reappraisal_total, data = df_final)
#summary(rq1)
#check_model(rq1)
```

```{r}

# RQ 2

## Is higher emotion suppression associated with higher depression symptoms in adolescent girls?

rq2<-lm(ces_dc_total_75perc~erq_suppression_total, data = df_final)
#summary(rq2)
#check_model(rq2)
```

```{r}

# Exploratory 

## Is higher suppression use associated with higher depression symptoms when adjusting for reappraisal use?

ex1 <- lm(ces_dc_total_75perc ~ erq_suppression_total + erq_reappraisal_total, data = df_final)
#summary(ex1)
#check_model(ex1)
```

```{r}

## Does age moderate the relation between reappraisal/suppression and depression symptoms in adolescent girls?

ex2 <- lm(ces_dc_total_75perc ~ erq_reappraisal_total * W1_mean_age, data = df_final)
#summary(ex2)
#check_model(ex2)

ex3 <- lm(ces_dc_total_75perc ~ erq_suppression_total * W1_mean_age, data = df_final)
#summary(ex3)
#check_model(ex3)
```

```{r}
# store info as objects for inline code below 
rq1_f <- summary(rq1)$fstatistic[1]
rq1_df2 <- summary(rq1)$fstatistic[3]
rq1_r2 <- summary(rq1)$r.squared
rq1_b <- coef(summary(rq1))["erq_reappraisal_total", "Estimate"]
rq1_se <- coef(summary(rq1))["erq_reappraisal_total", "Std. Error"]
rq1_t <- coef(summary(rq1))["erq_reappraisal_total", "t value"]

rq2_f <- summary(rq2)$fstatistic[1]
rq2_df2 <- summary(rq2)$fstatistic[3]
rq2_r2 <- summary(rq2)$r.squared
rq2_b <- coef(summary(rq2))["erq_suppression_total", "Estimate"]
rq2_se <- coef(summary(rq2))["erq_suppression_total", "Std. Error"]
rq2_t <- coef(summary(rq2))["erq_suppression_total", "t value"]


ex1_f  <- summary(ex1)$fstatistic[1]
ex1_df2 <- summary(ex1)$fstatistic[3]
ex1_r2 <- summary(ex1)$r.squared
ex1_b_reapp <- coef(summary(ex1))["erq_reappraisal_total", "Estimate"]
ex1_se_reapp <- coef(summary(ex1))["erq_reappraisal_total", "Std. Error"]
ex1_t_reapp <- coef(summary(ex1))["erq_reappraisal_total", "t value"]
ex1_b_supp <- coef(summary(ex1))["erq_suppression_total", "Estimate"]
ex1_se_supp <- coef(summary(ex1))["erq_suppression_total", "Std. Error"]
ex1_t_supp <- coef(summary(ex1))["erq_suppression_total", "t value"]

ex2_f <- summary(ex2)$fstatistic[1]
ex2_df2 <- summary(ex2)$fstatistic[3]
ex2_r2 <- summary(ex2)$r.squared
ex2_b_reapp <- coef(summary(ex2))["erq_reappraisal_total", "Estimate"]
ex2_se_reapp <- coef(summary(ex2))["erq_reappraisal_total", "Std. Error"]
ex2_t_reapp <- coef(summary(ex2))["erq_reappraisal_total", "t value"]
ex2_int_b <- coef(summary(ex2))["erq_reappraisal_total:W1_mean_age", "Estimate"]
ex2_int_se <- coef(summary(ex2))["erq_reappraisal_total:W1_mean_age", "Std. Error"]
ex2_int_t <- coef(summary(ex2))["erq_reappraisal_total:W1_mean_age", "t value"]

ex3_f <- summary(ex3)$fstatistic[1]
ex3_df2 <- summary(ex3)$fstatistic[3]
ex3_r2 <- summary(ex3)$r.squared
ex3_b_supp <- coef(summary(ex3))["erq_suppression_total", "Estimate"]
ex3_se_supp <- coef(summary(ex3))["erq_suppression_total", "Std. Error"]
ex3_t_supp <- coef(summary(ex3))["erq_suppression_total", "t value"]
ex3_int_b <- coef(summary(ex3))["erq_suppression_total:W1_mean_age", "Estimate"]
ex3_int_se <- coef(summary(ex3))["erq_suppression_total:W1_mean_age", "Std. Error"]
ex3_int_t <- coef(summary(ex3))["erq_suppression_total:W1_mean_age", "t value"]
```

## Emotion Regulation and Depression

The results of associations between emotion regulation strategies and depression are reported in Table 2. The association between cognitive reappraisal and depression is shown in Figure 1. A linear regression model was estimated predicting depressive symptoms from cognitive reappraisal. The overall model was significant, *F*(1, `r rq1_df2`) = `r round(rq1_f, 2)`, *p* \< .001, and accounted for approximately `r round(rq1_r2, 2)*100`% of the variance in depressive symptoms (*R*² = `r round(rq1_r2, 2)`). Cognitive reappraisal significantly predicted depressive symptoms, such that greater use of reappraisal was associated with lower symptom severity, *b* = `r round(rq1_b, 2)`, *SE* = `r round(rq1_se, 2)`, *t* = `r round(rq1_t, 2)`, *p* \< .001.

The association between expressive suppression and depression is shown in Figure 2. A linear regression was conducted to assess the relation between expressive suppression and depressive symptoms. The model was significant, *F*(1, `r rq2_df2`) = `r round(rq2_f, 2)`, *p* \< .001, accounting for about `r round(rq2_r2, 2)*100`% of the variance in depressive symptoms (*R*² = `r round(rq2_r2, 2)`). Expressive suppression significantly predicted depressive symptoms, with higher suppression associated with higher symptom scores, *b* = `r round(rq2_b, 2)`, *SE* = `r round(rq2_se, 2)`, *t* = `r round(rq2_t, 2)`, *p* \< .001.

### Table 2. Model Summary for RQ1 and RQ2

```{r}
## Table 2
modelsummary(
  list(
    "Reappraisal" = rq1,
    "Suppression" = rq2),
  coef_map = c(
    "(Intercept)"             = "Intercept",
    "erq_reappraisal_total"   = "Reappraisal (ERQ)",
    "erq_suppression_total"   = "Suppression (ERQ)"
  ),
  stars = TRUE
  )
```

### Figure 1

```{r}
## Plots

ggplot(df_final, aes(x=erq_reappraisal_total, y=ces_dc_total_75perc))+
  geom_point(color="Dodgerblue3")+
  geom_smooth(method = lm, color="Aquamarine3")+
  labs(x="Reappraisal", y="Depression", title = "Association between cognitive reappraisal and depressive symptoms") +
  theme_classic(base_size = 11)

```

### Figure 2

```{r}

ggplot(df_final,aes(x=erq_suppression_total, y=ces_dc_total_75perc))+
  geom_point(color="Dodgerblue3")+
  geom_smooth(method = lm, color="Aquamarine3")+
  labs(x="Suppression", y="Depression", title = "Association between expressive suppression and depressive sympotms") + 
  theme_classic(base_size = 11)

```

## Exploratory Analysis

The results of exploratory analysis are reported in Table 3. A multiple linear regression was conducted to examine whether expressive suppression and cognitive reappraisal independently predicted depressive symptoms. The overall model was significant, *F*(2, `r ex1_df2`) = `r round(ex1_f, 2)`, *p* \< .001, and explained approximately `r round(ex1_r2, 2)*100`% of the variance in depressive symptoms (*R*² = `r round(ex1_r2, 2)`). Expressive suppression emerged as a significant positive predictor, *b* = `r round(ex1_b_supp, 2)`, *SE* = `r round(ex1_se_supp, 2)`, *t* = `r round(ex1_t_supp, 2)`, *p* \< .001. Cognitive reappraisal was a significant negative predictor, *b* = `r round(ex1_b_reapp, 2)`, *SE* = `r round(ex1_se_reapp, 2)`, *t* = `r round(ex1_t_reapp, 2)`, *p* \< .001. These results suggest that suppression and reappraisal make independent and opposing contributions to depressive symptoms in early adolescent girls.

To test whether age moderated these associations, two interaction models were estimated. The interaction plots are shown in Figure 3. In the first model, cognitive reappraisal, age, and their interaction were entered as predictors. The overall model was significant, *F*(3, `r ex2_df2`) = `r round(ex2_f, 2)`, *p* \< .001, explaining `r round(ex2_r2, 2)*100`% of the variance. Greater reappraisal use predicted lower depressive symptoms, *b* = `r round(ex2_b_reapp, 2)`, *SE* = `r round(ex2_se_reapp, 2)`, *t* = `r round(ex2_t_reapp, 2)`, *p* \< .05, whereas age was not a significant predictor, *p* \> .05. The Reappraisal × Age interaction was not statistically significant, *b* = `r round(ex2_int_b, 2)`, *SE* = `r round(ex2_int_se, 2)`, *t* = `r round(ex2_int_t, 2)`, *p* \> .05, suggesting that the strength of the association between reappraisal and depressive symptoms did not differ by adolescent age.

A second moderation model examined whether age moderated the association between expressive suppression and depressive symptoms. The model was significant overall, *F*(3, `r ex3_df2`) = `r round(ex3_f, 2)`, *p* \< .001, accounting for `r round(ex3_r2, 2)*100`% of the variance. However, neither suppression nor age significantly predicted depressive symptoms, and the Suppression × Age interaction was non-significant (*p* \> .05), indicating that age did not moderate the relation between expressive suppression and depressive symptoms.

### Table 3. Model Summaries for Multiple Regression and Interaction Models

```{r}

# Table 3
modelsummary(
  list(
    "Suppression + Reappraisal" = ex1,
    "Reappraisal × Age Interaction" = ex2,
    "Suppression × Age Interaction" = ex3),
    coef_map = c(
    "(Intercept)"                     = "Intercept",
    "erq_reappraisal_total"   = "Reappraisal (ERQ)",
    "erq_suppression_total"   = "Suppression (ERQ)",
    "erq_reappraisal_total:W1_mean_age" = "Reappraisal x Age",
    "erq_suppression_total:W1_mean_age" = "Suppression x Age"),
  stars = TRUE
  )
```

### Figure 3

```{r}
# Interaction between Reappraisal and Depression
df_final2 <- df_final %>%
  mutate(
    age_group_2 = case_when(
      W1_mean_age >= 10 & W1_mean_age < 12 ~ "10-11",
      W1_mean_age >= 12 & W1_mean_age <= 13 ~ "12-13",
      TRUE ~ NA_character_
    ))

ggplot(
  df_final2%>%filter(!is.na(age_group_2)),
  aes(
    x=erq_reappraisal_total,
    y=ces_dc_total_75perc,
    linetype = age_group_2
  )
)+
  geom_smooth(
    method = "lm",
    se=TRUE,
    fullrange=TRUE,
    color="Aquamarine3",
    linewidth=1,
    alpha=0.15
  )+
  labs(
    x="Reappraisal",
    y="Depression",
    linetype="Age Group",
    title = "Interaction of Reappraisal and Age Predicting Depression Symptoms"
  )+
  scale_linetype_manual(values = c("solid","dashed"))+
  theme_classic(base_size = 11)+
  theme(
    legend.position = "bottom",
    legend.key = element_blank()
  )
```

### Figure 4

```{r}
#Suppression and Depression

ggplot(
  df_final2 %>% filter(!is.na(age_group_2)),
  aes(
    x = erq_suppression_total,
    y = ces_dc_total_75perc,
    linetype = age_group_2
  )
) +
  geom_smooth(
    method    = "lm",
    se        = TRUE,
    fullrange = TRUE,
    color     = "Aquamarine3",
    linewidth = 1,
    alpha     = 0.15
  ) +
  labs(
    x   = "Suppression",
    y   = "Depression",
    linetype = "Age Group",
    title = "Interaction of Suppression and Age Predicting Depression Symptoms"
  ) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_classic(base_size = 11) +
  theme(
    legend.position = "bottom",
    legend.key      = element_blank()
  )
```

# Discussion

The present study investigated associations between two core emotion regulation (ER) strategies, cognitive reappraisal and expressive suppression, and depressive symptoms among young adolescent girls, as well as whether age moderated these associations. Across analytic approaches, results revealed that greater use of cognitive reappraisal was linked to lower depressive symptoms, whereas greater use of expressive suppression was linked to higher depressive symptoms. These findings are consistent with the broader ER literature [@aldao2010; @compas2017] and demonstrate that these patterns are already evident during early adolescence, a developmental stage characterized by heightened emotional reactivity and growing vulnerability to internalizing disorders.

## Emotion Regulation as a Salient Predictor in Early Adolescence

Our findings underscore the relevance of ER strategy use for understanding depressive symptoms at a time when girls experience disproportionately high increases in depression risk [@rohde2009]. Reappraisal, a strategy involving reframing emotional situations to alter one’s emotional experience, was negatively associated with depressive symptoms, supporting its conceptualization as an adaptive and protective regulatory strategy [@dryman2018]. In contrast, suppression, a strategy aimed at inhibiting outward emotional expression, was associated with higher depressive symptoms, consistent with its categorization as a maladaptive strategy linked to emotional distancing, reduced social support, and increased physiological stress responses [@ehring2010].

Importantly, these associations persisted when the two strategies were modeled together, suggesting that each contributes uniquely to depressive symptoms rather than functioning as opposite ends of a single continuum. This distinction adds nuance to our understanding of ER processes that adolescents may simultaneously engage in both adaptive and maladaptive strategies, and the balance between them may contribute meaningfully to mental health.

## Developmental Considerations and Age as a Moderator

Exploratory analyses further revealed that age did not moderate the association between reappraisal and depressive symptoms. Although reappraisal was associated with lower depressive symptoms across the full sample, this protective association appeared weaker among older adolescents; however, this result was not statistically significant. Thus, it is possible that that developmental changes in emotional, cognitive, or social processes may influence the effectiveness with which reappraisal is deployed across the early and later stages of adolescence; however, more research is needed to investigate this potential moderation effect, particularly across a larger range of adolescents. For instance, younger adolescents may benefit more from cognitive reframing strategies as they begin to encounter more complex emotional challenges. In contrast, older adolescents may increasingly rely on a broader set of regulatory skills or face social contexts that diminish the effectiveness of reappraisal. This finding is consistent with conceptual models that propose early adolescence as a sensitive period for the development of ER skills [@silvers2022] and underscores the importance of strengthening adaptive regulation strategies before adolescence progresses further.

Moreover, age did not moderate the association between expressive suppression and depressive symptoms. Suppression was linked to elevated depressive symptoms regardless of age. This suggests that suppression may function as a consistently maladaptive strategy across early adolescence, possibly because it interferes with emotional expression, social connectedness, and psychological flexibility in ways that are detrimental at any age during this developmental period.

## Implications for Theory, Prevention, and Intervention

These findings have several theoretical and applied implications. Theoretically, they support models positioning ER as a transdiagnostic factor in the development of internalizing symptoms [@aldao2016]. They also align with neurodevelopmental models suggesting that early adolescence marks a sensitive period for the maturation of prefrontal and limbic circuitry involved in cognitive regulation of emotion [@gee2016]. Because reappraisal relies heavily on prefrontal control processes, its benefits may be particularly meaningful when these neural systems are still maturing.

From a prevention and intervention standpoint, our results highlight reappraisal as a potential target for early skill-building efforts. Programs that teach adolescents how to reinterpret emotional situations, such as cognitive-behavioral interventions, may be especially effective if delivered early, before the benefits of reappraisal begin to diminish. Similarly, interventions that reduce reliance on expressive suppression and promote more adaptive coping strategies may mitigate risk for depression onset. Importantly, the moderation findings suggest that the timing of the intervention matters. Given that the protective effect of reappraisal appears stronger in the earlier years of adolescence, prevention programs may benefit from being implemented around ages 10-12, when cognitive ER skills may be more malleable and impactful.

## Limitations and Future Directions

Several limitations should be considered when interpreting these findings. First, the present study relied on baseline cross-sectional data, which limits our ability to draw conclusions about directionality or causal pathways between ER and depressive symptoms. Longitudinal analyses are needed to determine whether ER strategies predict changes in depressive symptoms over time, or vice versa. Second, all variables were assessed via self-report, which may introduce shared method variance or reporting biases. Incorporating multi-method assessments, including parent-report, behavioral tasks, or physiological indicators, would help strengthen the validity of future findings. Third, although our sample was demographically representative of the geographic region, it was not highly diverse, and replication in more heterogeneous samples would improve generalizability. Finally, our models focused on two ER strategies, but adolescents use a vast repertoire of regulation behaviors; future work should examine additional strategies and consider how they interact within broader ER profiles.

## Conclusion

Overall, the present study provides evidence that cognitive reappraisal and expressive suppression are meaningful correlates of depressive symptoms in early adolescent girls. Reappraisal appears to confer a protective effect, whereas suppression is associated with greater depressive symptoms. The moderating role of age suggests that the benefits of reappraisal may diminish as adolescents grow older, underscoring the importance of strengthening adaptive ER skills during early adolescence. These findings lay the groundwork for future longitudinal work aimed at understanding how ER strategies shape developmental trajectories of depression and identifying early targets for intervention and prevention

# References
