---
title: "Final Project"
authors: "Xinyu (Lily) Wang, Sophie Hartford, Everett Mahaffy"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    embed-resources: true
    highlight-style: tango
execute:
  eval: true
  echo: true
warning: false
message: false
---

# Load packages

```{r}
library(tidyverse)
library(flextable)
library(here)
library(readxl)
library(modelsummary)
library(janitor)
library(performance)
```

# Load data

```{r}
# Demographic
age <- read_csv(here("data", "TAG_age_session_dates_4waves.csv"))
age <- clean_names(age)

race <- read_excel(here("data", "TAG_W1_Race_Ethnicity.xlsx"))
race <- clean_names(race)

# Emotion Regulation(ERQ)
erq <- read_csv(here("data", "ERQ_Wave1.csv"))
erq <- erq %>% 
       clean_names() %>% 
       distinct(tagid, .keep_all = TRUE)

# Adolescent Depression(CESDC)
cesdc <- read_csv(here("data", "CESDC_Wave1.csv"))
cesdc <- clean_names(cesdc)

```

# Participant demographics
## Race/ethnicity category
### Clean the data frame: identify subjects reported multiple racial categories and rename them to "Multi-racial"

```{r}
race_clean <- race %>%
  mutate(
    w1_ethnicity = as.character(w1_ethnicity),
    w1_ethnicity = if_else(
      str_detect(w1_ethnicity, ","),
      "g. Multi-racial",
      w1_ethnicity
    )
  )
```

# Age
## We have subject age data at both sessions and we are interested in calculating the mean age of the two sessions of wave1. 

```{r}
age <- age %>% 
  mutate(W1_mean_age = (w1s1_age + w1s2_age)/2) 

# Check mean age for sample
age_summary <- age %>%
  summarize(
    mean_W1_age = mean(W1_mean_age, na.rm = TRUE),
    sd_W1_age   = sd(W1_mean_age, na.rm = TRUE),
  )
print(age_summary)
```

# Join all data frames

```{r}
df <- erq %>% 
  left_join(cesdc, by= 'tagid')

df <- df %>% 
     left_join(age, by = 'tagid')

df <- df %>% 
     left_join(race_clean, by = 'tagid') %>% 
     select(tagid, erq_reappraisal_total, erq_suppression_total, ces_dc_total_75perc, W1_mean_age, w1_ethnicity, 
            w1s1_date, w1s1_age, w1s2_date, w1s2_age, w2s1_date, w2s1_age, w2s2_date, w2s2_age, 
            w3s1_date, w3s1_age, w3s2_date, w3s2_age, w4s1_date, w4s1_age, w4s2_date, w4s2_age
            )

df <- df %>%
  rename(
    wave1_session1_date = w1s1_date,
    wave1_session1_age  = w1s1_age,
    wave1_session2_date = w1s2_date,
    wave1_session2_age  = w1s2_age,

    wave2_session1_date = w2s1_date,
    wave2_session1_age  = w2s1_age,
    wave2_session2_date = w2s2_date,
    wave2_session2_age  = w2s2_age,

    wave3_session1_date = w3s1_date,
    wave3_session1_age  = w3s1_age,
    wave3_session2_date = w3s2_date,
    wave3_session2_age  = w3s2_age,

    wave4_session1_date = w4s1_date,
    wave4_session1_age  = w4s1_age,
    wave4_session2_date = w4s2_date,
    wave4_session2_age  = w4s2_age
  )
```

# Tidy and filter

```{r}
# Tidy
df2 <- df %>%
  mutate(across(starts_with("wave"), as.character))

df_long <- df2 %>%
  pivot_longer(
    cols = starts_with("wave"),
    names_to = c("wave", "session", "info"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  mutate(
    wave    = parse_number(wave),
    session = parse_number(session)
  )

df_tidy <- df_long %>%
  pivot_wider(
    names_from = info,
    values_from = value
  )

# Filter
df_tidy <- df_tidy %>% 
  filter(wave == 1, session == 1)
```

# Sample Summary Statistics

```{r}
df_names <- df_tidy %>% 
    select(-wave, -session) %>% 
    rename(
    Reappraisal       = erq_reappraisal_total,
    Suppression       = erq_suppression_total,
    Depression        = ces_dc_total_75perc,
    Age               = W1_mean_age,
    Ethnicity         = w1_ethnicity
    )
datasummary_skim(df_names,
                 title = "Table 1. Sample Descriptive Statistics")
```

# RQ 1

## Is higher emotion reappraisal associated with lower depression symptoms in adolescent girls?

```{r}
rq1 <- lm(df_tidy$ces_dc_total_75perc ~ df_tidy$erq_reappraisal_total)
check_model(rq1)
```

# RQ 2

## Is higher emotion suppression associated with higher depression symptoms in adolescent girls?

```{r}
rq2 <- lm(df_tidy$ces_dc_total_75perc ~ df_tidy$erq_suppression_total)
check_model(rq2)
```

# Exploratory 

## Is higher suppression use associated with higher depression symptoms when adjusting for reappraisal use?

```{r}
ex1 <- lm(df_tidy$ces_dc_total_75perc ~ df_tidy$erq_suppression_total + df_tidy$erq_reappraisal_total)
check_model(ex1)
```

## Does age moderate the relation between reappraisal/suppression and depression symptoms in adolescent girls?

```{r}
ex2 <- lm(df_tidy$ces_dc_total_75perc ~ df_tidy$erq_reappraisal_total * df_tidy$W1_mean_age)
check_model(ex2)

ex3 <- lm(df_tidy$ces_dc_total_75perc ~ df_tidy$erq_suppression_total * df_tidy$W1_mean_age)
check_model(ex3)
```

# Results

## Tables
```{r}
modelsummary(
  list(
    "Reappraisal" = rq1,
    "Suppression" = rq2),
  coef_map = c(
    "(Intercept)"                     = "Intercept",
    "df_tidy$erq_reappraisal_total"   = "Reappraisal (ERQ)",
    "df_tidy$erq_suppression_total"   = "Suppression (ERQ)"
  ),
  stars = TRUE,
  title = "Table 2. Model Summary for RQ1 and RQ2"
  )

modelsummary(
  list(
    "Suppression + Reappraisal" = ex1,
    "Reappraisal × Age Interaction" = ex2,
    "Suppression × Age Interaction" = ex3),
    coef_map = c(
    "(Intercept)"                     = "Intercept",
    "df_tidy$erq_reappraisal_total"   = "Reappraisal (ERQ)",
    "df_tidy$erq_suppression_total"   = "Suppression (ERQ)",
    "df_tidy$erq_reappraisal_total:df_tidy$W1_mean_age" = "Reappraisal x Age",
    "df_tidy$erq_suppression_total:df_tidy$W1_mean_age" = "Suppression x Age"),
  stars = TRUE,
  title = "Table 3. Model Summaries for Multiple Regression and Interaction Models"
  )
```

## Plots
```{r}
## two scatter plots with regression lines (method lm) here: one for reappraisal (x axis) and one for suppression (x axis); y axis will be depression for both
```

# Discussion
## [ to do later ]